Computing likelihoods under the SMC for a general class of ARGs.

### 2024-07-05 paper.tex
A pass at the abstract

\begin{abstract}
There currently is no way to connect the output of the diverse
range of ARG inference methods back
to a model that would allow us to perform a probabilistic exploration of the
ARG space around the point estimate, or likelihood-based inference.
\end{abstract}

>>>>

\begin{abstract}
Ancestral recombination graphs (ARGs) are the focus of much ongoing research
interest. Recent progress in inference has made ARG-based approaches feasible
across of range of applications, and many new methods using inferred ARGs as
input have appeared. This progress on the long-standing problem of ARG
inference has proceeded in two distinct directions. The first
class of method in which dramatic improvements have been made is in the
statistically rigorous inference of ARGs under the Sequentially Markov
Coalescent (SMC) model, which are now practical for tens-to-hundreds of
samples. The second class of recently-developed inference methods are based on
approximate models and heuristics which scale to samples sizes roughly three
orders of magnitude larger. Although these heuristic methods are reasonably
accurate under many metrics, one major problem is that the ARGs they 
estimate cannot easily be connected with population genetic theory.
Specifically, because the estimated ARGs lack detail about recombination
events, there is currently no way to compute their likelihood
under a population genetics model. In this paper we provide a 
backwards-time formulation of the SMC based on a total ordering of 
lineages, and show how this leads to straightforward description 
of the likelihood of a general class of ARG under the SMC. We show that this 
formulation does not require precise details of recombination events
to be estimated, and is robust to other topological issues such as 
polytomies. This opens up many future directions, most excitingly the possibility
of combining the strengths of both inference approaches. Applying
these ideas may allow us to 
make rigorous inference more scalable, and scalable inference 
more rigorous.
\end{abstract}
